### 2022-09-28 main/chapter_theory.tex
add mssm mass sentence
cut out preface junk; fix more atlas's; minor tweaks

\begin{singlespacing}
\chapter{The Standard Model and beyond}
\label{chapter:theory}
%
\begin{epigraphs}
\qitem{%
if observing outer space gives us a view of the past, observing inner space
would surely give us a glimpse into the future - would be interesting if NASA
made a telescope for that%
}%
{Ken~M,
\textit{Comment: Yahoo! News},
2012~\cite{kenm2012inner}}
\end{epigraphs}
\end{singlespacing}

\noindent Our best current understanding of nature's fundamental objects in
their observable behaviours as particles is encoded in the Standard Model,
which uses solid mathematical foundations to accurately describe the physics
of particle interactions across a broad domain.
But that domain is bounded, so the Standard Model is not an ultimate theory of
everything, and work continues to expand that boundary on both theoretical and
experimental fronts.
The current Standard Model is described in Section~\ref{sec:context_sm},
along with its limitations and possible future directions.

Supersymmetry an idea that has been applied to build plausible alternatives
to the Standard Model which may explain future experimental discoveries.
Since the main work of this thesis is a search for effects of supersymmetric
phenomena, supersymmetry and the specific supersymmetric models that we study
are presented in Section~\ref{sec:context_susy}.

\section{The Standard Model}
\label{sec:context_sm}
The current Standard Model features fermions
(of four types in three generations),
which interact through forces mediated by vector bosons
and the scalar Higgs boson.
The strong force is mediated by
Massless gluons $g$ mediate the strong force,
the massless photon $\gamma$ mediates the electromagnetic force,
and very massive $W^+$, $W^-$, and $Z$ bosons mediate the weak force.
These fundamental particles of the Standard Model are summarized in
Table~\ref{tab:context_particles_sm}.

@@

% using https://www.tablesgenerator.com/latex_tables
\begin{table}[tp]
\centering
\begin{tabular}{rrc}
\multirow{4}{*}{fermions} & \multirow{2}{*}{quarks}      & \makebox[5em]{\hfill $u$ \hfill $c$ \hfill $t$ \hfill}                    \\
                          &                              & \makebox[5em]{\hfill $d$ \hfill $s$ \hfill $b$ \hfill}                    \\[1ex]
                          & \multirow{2}{*}{leptions}    & \makebox[5em]{\hfill $e$ \hfill $\mu$ \hfill $\tau$ \hfill}               \\
                          &                              & \makebox[5em]{\hfill $\nu_e$ \hfill $\nu_\mu$ \hfill $\nu_\tau$ \hfill} \\[3ex]
\multirow{2}{*}{bosons}   & strong                       & \makebox[7em]{\hfill $g$ \hfill}                                          \\[1ex]
                          & electroweak & \makebox[7em]{\hfill $W^\pm\!\!$ \hfill $Z$\hfill $\gamma$ \hfill $h$ \hfill} \\
\end{tabular}
\caption[Particles of the Standard Model]{%
Particles of the Standard Model.
All fermions are spin-$1/2$, and all bosons are spin-$1$ except for the
spin-$0$ Higgs boson $h$.
Fermions come in left- and right-chiralities, although the existence of
right-chiral neutrinos is not strictly confirmed.%
}
\label{tab:context_particles_sm}
\end{table}

@@

\paragraph{Neutrino masses.}
Neutrinos could be Dirac fermions with Yukawa masses and boring right-chiral
fields.
They could also have Majorana mass terms, for which the seesaw mechanism can
simultaneously generate the observed light neutrinos with very heavy `sterile'
neutrinos.
Or their masses could come from other operators without needing the
right-chiral fields~\cite{
thomson2013modern,
chala2021neutrino,
wells2020discovery
}.
This answer is uncertain, and new results from theory or experiment that
pinpoint a correct model could improve the Standard Model in the neutrino
domain.

\paragraph{Gravitation and cosmology.}
Gravity does not feature in the Standard Model, nor does it measurably impact
particle physics experiments.
Although gravitation is in another domain, the particle content of the
Standard Model does impact cosmology.

@@

\paragraph{Origins.}
Why does the Standard Model have its structure?
Why is it configured with the observed values of our parameters?
Why do we even live in $3+1$ dimensional spacetime with quantum mechanics?
Answers to these questions are largely unknown, but would be
attractive features for a fundamental theory.

@@

\paragraph{Aesthetics.}
In the search for improved models, heuristic criteria can help to identify
appealing candidates.
Two such criteria are `fine-tuning' and `naturalness', both of which are
aesthetic judgements about dimensionless model parameters.
Naturalness is opinion; only nature decides what is really natural,
but these criteria can still guide the discourse of our scientific searches
towards useful directions~\cite{wells2020discovery}.

@@

\section{Supersymmetric models}
\label{sec:context_susy}
Supersymmetry is the concept of symmetry under the operation of exchanging
bosonic and fermionic states, and supersymmetric models use supersymmetry in
their construction.
Nature is not totally supersymmetric.
If it were, then there would be bosonic `superpartners' for each of the known
fermions (and vice versa) with the same masses and charges, and those
`sparticles' would have been seen long ago.
But supersymmetry might still participate in a broken condition if the
sparticles are hidden at larger masses.

Any new model must include the observational facts of the Standard Model,
and the Minimal Supersymmetric Standard Model (MSSM) satisfies this while
adding a school of new particles,
which is listed in Table~\ref{tab:context_particles_mssm}.

% fermion boson multiplets
The MSSM matches each standard fermion to a pair of scalar
`sfermions' (left and right)
and each vector gauge field to a `gaugino' fermion
($\tilde g$, $\tilde W^1$, $\tilde W^2$, $\tilde W^3$, and $\tilde B$).
The Higgs field does not translate so directly, but is replaced with two sets
of complex scalar fields
($H^+_u$, $H^0_u$ and $H^0_d$, $H^\pm_d$) with
gaugino superpartners
($\tilde H^+_u$, $\tilde H^0_u$ and $\tilde H^0_d$, $\tilde H^-_d$).
(All sparticles are marked with tildes $\tilde{~}$.)
These fields mix into mass eigenstates which current data do not entirely
determine:
the gaugino mixtures are the charged charginos $C^\pm_i$ and
neutral neutralinos $N_i$
(also labelled $\chargino_i$ and $\neutralino_i$~\cite{atlas2022searches}),
and each squark and slepton has two states that mix the left and right fields.
From the four complex scalars, three degrees of freedom are eaten by the
weak bosons to leave the reborn Standard Model Higgs boson $h$
and four new scalars~\cite{
martin2016primer
}.

% made using:
% https://www.tablesgenerator.com/latex_tables
\begin{table}[tp]
\centering
\begin{tabular}{rrc}
\multirow{5}{*}{bosons}   & Higgs                         & \makebox[7em]{\hfill ($h$) \hfill $H^0$ \hfill $A^0$ \hfill $H^\pm\!\!$ \hfill}             \\[1ex]
                          & \multirow{2}{*}{squarks}      & \makebox[5em]{\hfill $\tilde u$ \hfill $\tilde c$ \hfill $\tilde t$ \hfill}                  \\
                          &                               & \makebox[5em]{\hfill $\tilde d$ \hfill $\tilde s$ \hfill $\tilde b$ \hfill}                  \\[1ex]
                          & \multirow{2}{*}{sleptions}    & \makebox[5em]{\hfill $\tilde e$ \hfill $\tilde \mu$ \hfill $\tilde \tau$ \hfill}             \\
                          &                               & \makebox[5em]{\hfill $\tilde \nu_e$ \hfill $\tilde \nu_\mu$ \hfill $\tilde \nu_\tau$ \hfill} \\[3ex]
\multirow{3}{*}{fermions} & gluinos                       & \makebox[5em]{\hfill $\tilde g$ \hfill}                                        \\[1ex]
                          & \multirow{2}{*}{gauginos}     & \makebox[5em]{\hfill $\tilde C^\pm_1\!\!$ \hfill $\tilde C^\pm_2\!\!$ \hfill} \\
                          &                               & \makebox[7em]{\hfill $\tilde N_1\!$ \hfill $\tilde N_2\!$ \hfill$\tilde N_3\!$ \hfill$\tilde N_4\!$ \hfill} \\
\end{tabular}
\caption[%
Particles of the Minimal Supersymmetric Standard Model%
]{%
Particles of the Minimal Supersymmetric Standard Model
in their mass eigenstates~\cite{martin2016primer}, in addition to the
Standard Model particles of Table~\ref{tab:context_particles_sm}.
The standard Higgs boson ($h$) is repeated here with its new pals.
Sparticles are marked with tildes $\tilde{~}$.
All listed bosons are spin-$0$ and all listed fermions are spin-$1/2$.
}
\label{tab:context_particles_mssm}
\end{table}

@@

Supersymmetric modelling successfully predicted the lightest Higgs boson
mass to be below $\lesssim 130\,\eV[G]$ long before its observation at
$125\,\eV[G]$~\cite{
espinosa1992higgs,
espinosa1993upper,
quiros1997constraints,
wells2020discovery,
atlas2012higgs,
atlas2012combined,
cms2012higgs
}.
This prediction, however, assumed a mass scale below $1\,\eV[T]$,
and observations at the LHC are now limiting sparticle masses
into super-$\eV[T]$ territory (in specific models).
Such limits are also troublesome for the hierarchy problem, because large
superpartner masses introduce fine-tuning problems of their
own~\cite{martin2016primer}.
Despite its current lack of observational evidence, however,
supersymmetry remains a tool for building theoretically plausible
models, and might help us to understand future discoveries.

>>>>

\begin{singlespacing}
\chapter{The Standard Model and beyond}
\label{chapter:theory}
%
\begin{epigraphs}
\qitem{%
The idea that theorems follow from the postulates does not correspond to simple
observation.
If the Pythagorean theorem were found to not follow from the postulates, we
would again search for a way to alter the postulates until it was true.
Euclid's postulates came from the Pythagorean theorem, not the other way.}%
{Richard~Hamming,
\textit{The Unreasonable Effectiveness of Mathematics},
1980~\cite{hamming1980unreasonable}}
\end{epigraphs}
\end{singlespacing}
\noindent
The Standard Model encodes out best understanding of nature's fundamental
objects in their behaviours as particles.
From solid mathematical foundations, it accurately describe the physics
of particle interactions across a broad domain,
but that domain is bounded, so the Standard Model is not an ultimate theory of
everything.
Work continues to expand that boundary on both theoretical and experimental
fronts.
The current Standard Model is described in Section~\ref{sec:theory_sm},
along with its limitations and possible future directions.

Supersymmetry an idea that has been applied to build plausible alternatives
to the Standard Model which may explain future experimental discoveries.
Since the main work of this thesis is a search for effects of supersymmetric
phenomena, supersymmetry and the specific supersymmetric models that we study
are presented in Section~\ref{sec:theory_susy}.

\section{The Standard Model}
\label{sec:theory_sm}
The current Standard Model features fermions
(of four types in three generations),
which interact through forces mediated by vector bosons
and the scalar Higgs boson.
The strong force is mediated by
Massless gluons $g$ mediate the strong force,
the massless photon $\gamma$ mediates the electromagnetic force,
and very massive $W^+$, $W^-$, and $Z$ bosons mediate the weak force.
These fundamental particles of the Standard Model are summarized in
Table~\ref{tab:theory_particles_sm}.

@@

% using https://www.tablesgenerator.com/latex_tables
\begin{table}[tp]
\centering
\begin{tabular}{rrc}
\multirow{4}{*}{fermions} & \multirow{2}{*}{quarks}      & \makebox[5em]{\hfill $u$ \hfill $c$ \hfill $t$ \hfill}                    \\
                          &                              & \makebox[5em]{\hfill $d$ \hfill $s$ \hfill $b$ \hfill}                    \\[1ex]
                          & \multirow{2}{*}{leptions}    & \makebox[5em]{\hfill $e$ \hfill $\mu$ \hfill $\tau$ \hfill}               \\
                          &                              & \makebox[5em]{\hfill $\nu_e$ \hfill $\nu_\mu$ \hfill $\nu_\tau$ \hfill} \\[3ex]
\multirow{2}{*}{bosons}   & strong                       & \makebox[7em]{\hfill $g$ \hfill}                                          \\[1ex]
                          & electroweak & \makebox[7em]{\hfill $W^\pm\!\!$ \hfill $Z$\hfill $\gamma$ \hfill $h$ \hfill} \\
\end{tabular}
\caption[Particles of the Standard Model]{%
Particles of the Standard Model.
All fermions are spin-$1/2$, and all bosons are spin-$1$ except for the
spin-$0$ Higgs boson $h$.
Fermions come in left- and right-chiralities, although the existence of
right-chiral neutrinos is not strictly confirmed.%
}
\label{tab:theory_particles_sm}
\end{table}

@@

\subsection{Neutrino masses}
Neutrinos could be Dirac fermions with Yukawa masses and boring right-chiral
fields.
They could also have Majorana mass terms, for which the seesaw mechanism can
simultaneously generate the observed light neutrinos with very heavy `sterile'
neutrinos.
Or their masses could come from other operators necessarily introducing the
right-chiral fields~\cite{
thomson2013modern,
chala2021neutrino,
wells2020discovery
}.
The answer is uncertain, and new results from theory or experiment that
pinpoint a correct model could improve the Standard Model in the neutrino
domain.

\subsection{Gravitation and cosmology}
Gravity does not feature in the Standard Model, nor does it measurably impact
particle physics experiments.
Although gravitation is in another domain, the particle content of the
Standard Model does impact cosmology.

@@

\subsection{Origins}
Why does the Standard Model have its structure?
Why is it configured with the observed values of our parameters?
Why do we even live in $3+1$ dimensional spacetime with quantum mechanics?
Answers to these questions are largely unknown, but would be
attractive features for a fundamental theory.

@@

\subsection{Aesthetics}
In the search for improved models, heuristic criteria can help to identify
appealing candidates.
Two such criteria are `fine-tuning' and `naturalness', both of which are
aesthetic judgements about dimensionless model parameters.
Naturalness is opinion; only nature decides what is really natural,
but these criteria can still guide the discourse of our scientific searches
towards useful directions~\cite{wells2020discovery}.

@@

\section{Supersymmetric models}
\label{sec:theory_susy}
Supersymmetry is the concept of symmetry under the operation of exchanging
bosonic and fermionic states, and supersymmetric models use supersymmetry in
their construction.
Nature is not totally supersymmetric.
If it were, then there would be bosonic `superpartners' for each of the known
fermions (and vice versa) with the same masses and charges, and those
`sparticles' would have been seen long ago.
But supersymmetry might still participate in a broken condition if the
sparticles are hidden at larger masses.

Any new model must include the observational facts of the Standard Model,
and the Minimal Supersymmetric Standard Model (MSSM) satisfies this while
adding a school of new particles,
which is listed in Table~\ref{tab:theory_particles_mssm}.

The MSSM matches each standard fermion to a pair of scalar
`sfermions' (left and right)
and each vector gauge field to a `gaugino' fermion
($\tilde g$, $\tilde W^1$, $\tilde W^2$, $\tilde W^3$, and $\tilde B$).
The Higgs field does not translate so directly, but is replaced with two sets
of complex scalar fields
($H^+_u$, $H^0_u$ and $H^0_d$, $H^\pm_d$) with
gaugino superpartners
($\tilde H^+_u$, $\tilde H^0_u$ and $\tilde H^0_d$, $\tilde H^-_d$).
(All sparticles are marked with tildes $\tilde{~}$.)
These fields mix into mass eigenstates which current data do not entirely
determine:
the gaugino mixtures are the charged charginos $C^\pm_i$ and
neutral neutralinos $N_i$
(also labelled $\chargino_i$ and $\neutralino_i$~\cite{atlas2022searches}),
and each squark and slepton has two states that mix the left and right fields.
From the four complex scalars, three degrees of freedom are eaten by the
weak bosons to leave the reborn Standard Model Higgs boson $h$
and four new scalars~\cite{
martin2016primer
}.

Current observations do limit the masses and mixings of this new
school of MSSM particles, but their values remain very uncertain and in need
of new experimental constraints.

% made using:
% https://www.tablesgenerator.com/latex_tables
\begin{table}[tp]
\centering
\begin{tabular}{rrc}
\multirow{5}{*}{bosons}   & Higgs                         & \makebox[7em]{\hfill ($h$) \hfill $H^0$ \hfill $A^0$ \hfill $H^\pm\!\!$ \hfill}             \\[1ex]
                          & \multirow{2}{*}{squarks}      & \makebox[5em]{\hfill $\tilde u$ \hfill $\tilde c$ \hfill $\tilde t$ \hfill}                  \\
                          &                               & \makebox[5em]{\hfill $\tilde d$ \hfill $\tilde s$ \hfill $\tilde b$ \hfill}                  \\[1ex]
                          & \multirow{2}{*}{sleptions}    & \makebox[5em]{\hfill $\tilde e$ \hfill $\tilde \mu$ \hfill $\tilde \tau$ \hfill}             \\
                          &                               & \makebox[5em]{\hfill $\tilde \nu_e$ \hfill $\tilde \nu_\mu$ \hfill $\tilde \nu_\tau$ \hfill} \\[3ex]
\multirow{3}{*}{fermions} & gluinos                       & \makebox[5em]{\hfill $\tilde g$ \hfill}                                        \\[1ex]
                          & \multirow{2}{*}{gauginos}     & \makebox[5em]{\hfill $\tilde C^\pm_1\!\!$ \hfill $\tilde C^\pm_2\!\!$ \hfill} \\
                          &                               & \makebox[7em]{\hfill $\tilde N_1\!$ \hfill $\tilde N_2\!$ \hfill$\tilde N_3\!$ \hfill$\tilde N_4\!$ \hfill} \\
\end{tabular}
\caption[%
Particles of the Minimal Supersymmetric Standard Model%
]{%
Particles of the Minimal Supersymmetric Standard Model
in their mass eigenstates~\cite{martin2016primer}, in addition to the
Standard Model particles of Table~\ref{tab:theory_particles_sm}.
The standard Higgs boson ($h$) is repeated here with its new pals.
Sparticles are marked with tildes $\tilde{~}$.
All listed bosons are spin-$0$ and all listed fermions are spin-$1/2$.
}
\label{tab:theory_particles_mssm}
\end{table}

@@

Supersymmetric modelling successfully predicted the lightest Higgs boson
mass to be $\lesssim 130\,\eV[G]$ long before its observation at
$125\,\eV[G]$~\cite{
espinosa1992higgs,
espinosa1993upper,
quiros1997constraints,
wells2020discovery,
atlas2012higgs,
atlas2012combined,
cms2012higgs
}.
This prediction, however, assumed a mass scale below $1\,\eV[T]$,
and observations at the LHC are now limiting sparticle masses
into super-$\eV[T]$ territory (in specific models).
Such limits are also troublesome for the hierarchy problem, because large
superpartner masses introduce fine-tuning problems of their
own~\cite{martin2016primer}.
Despite its current lack of observational evidence, however,
supersymmetry remains a tool for building theoretically plausible
models, and might help us to understand future discoveries.

### 2022-09-28 main/draft.tex
cut out preface junk; fix more atlas's; minor tweaks

>>>>

% far too much intro waffle

\TODO{This does not belong in an introduction and will be moved.}

\TODO{CUT THIS SHORT}

\subsection{Testing symmetries in data}
Data-driven tests of physical symmetries are a speciality of the Cambridge
\atlas\ group.
Each test considers a theoretically motivated symmetry that states that
distributions of data should be identical when transformed under its operation.
This means that it can be possible to disprove a symmetry by searching only
for differences between data and transformed versions of those data, without
necessarily requiring other theoretical inputs.

% emus
Charge symmetry of electron-muon, `emu', pairs is a key example among these
tests.
Theoretically, it proposes that, in the absence of biases, data
from proton-proton collisions at the LHC would include equal distributions
of $e^+\mu^-$ and $e^-\mu^+$ events~\cite{Lester:2016qdv}.
Real data can have emu biases --- one clear (but minor) example is that our
detectors are made of matter containing negatively charged electrons, not
positrons, so high energy electrons positrons have subtly different rates of
energy loss in that detector matter.
However, \cite{Lester:2016qdv} argues that any biases at the LHC should
conspire only to favour the rate of $e^-\mu^+$ over the rate of $e^+\mu^-$
events.
Therefore defining an asymmetry
\begin{equation}
A^\mathrm{emu} = \frac{
\sigma(e^+\mu^- + X)
}{
\sigma(e^-\mu^+ + X)
}
,
\end{equation}
one can predict $A^\mathrm{emu} \leq 1$ from the Standard Model at the LHC.
Negative values could be explained by experimental biases, but
$A^\mathrm{emu} > 1$ measured in data might indicate new physics, since it and
can be caused by models of new physics that might invoke R-parity violating
supersymmetry~\cite{Lester:2016qdv} or leptoquarks~\cite{EXOT-2018-29}.
After its extensive development~\cite{Brunt:2674708,Pacey:2747774},
this test has been performed in \atlas\ data, and found no ``significant
model-independent evidence''~\cite{EXOT-2018-29} for $A^\mathrm{emu} > 1$.

% parity violation
Parity symmetry, which changes the signs of all spatial coordinates
$\vec x\rightarrow -\vec x$, is a second key example.
Nature does violate parity symmetry, as first shown by the Wu
experiment~\cite{PhysRev.105.1413}, in the weak sector, but that is not
directly observable at the LHC due to its non-polarized beams and
polarization-insensitive detectors~\cite{Lester:2019bso}.
Unknown physical effects could, however, generate forms of parity asymmetry
that could be seen at the LHC.
Although the charge symmetry of the emu test has known sources of bias,
no such biases are so clearly expected for parity.

General parity symmetry can therefore be tested at the LHC by considering a
parity-odd event variable and comparing its distribution to its parity-flipped
version, which simply swaps its sign.
% alpha from lester schott
One such test has been performed with LHC data~\cite{Lester:2019bso}.
This test used \cms\ open data and the parity-odd event variable
\begin{equation}
\label{eqn:cms_alpha}
\alpha
= \arcsin\!
\left(
\frac{
\vec p^{\,j_1} \times \vec p^{\,j_2}
}{
|\vec p^{\,j_1} \times \vec p^{\,j_2}|
}
\cdot
\frac{
\vec p^{\,j_3}
}{
|\vec p^{\,j_3}|
}
\right)
,
\end{equation}
in which $p^{\,j_i}$ is the momentum of the $i$th hardest jet in a multi-jet
event,
and did not find evidence for parity violation.
But that does not prove that no parity-violating effects are occuring in these
data, only that it did not appear in the distribution of $\alpha$.

Can a single parity-odd event variable be sensitive to all norms of parity
violation?
The answer, it turns out, is `no' if that variable is also required to be
continuous~\cite{lesterChiralMeasurements2021}.
In this framework, therefore, a joint analysis of multiple variables is
required for any sufficiently complicated data.
Furthermore, it is not generally easy to construct such universally-sensitive
sets of event variables~\cite{
Gripaios:2020hya,
lester2021lorentz
}.

Rather than beginning from universally sensitive event variables, an
alternative strategy is to build candidate variables to the data
themselves~\cite{lester2021stressed}.
If done wrong, such an idea could simply reinforce variations due to simple
statistical noise.
It can, however, be formulated as a rigorous test if one employs standard
Machine Learning practice and splits the data into independent training and
testing sets.
Variables learned (by function approximation algorithms such as
Boosted Decision Trees~\cite{xgboost} or
Neural Networks~\cite{MurphyKevinP.2012Mlap})
from the testing set can correspond to hypotheses about possible asymmetries
in the data distribution, and if those hypotheses make accurate predictions in
independent testing data, then they are evidence against the symmetry.

My contribution to this field is to adapt this idea of learning parity
variables,
\emph{which was initiated by Christopher~G.~Lester}~\cite{lester2021stressed},
into a probabilistic description, in which the algorithms predict
`which is real?' between real data an their parity-transformed
copies~\cite{tombs2021which}.
This approach derives an objective function
(which is a special case of the binary
cross-entropy~\cite{MurphyKevinP.2012Mlap}),
from which learning algorithms
can approximate the odds ratio between the data distribution and its
parity-transformed version.
Importantly, it also allows one to model inefficiencies or holes in the
distribution, which exist in practice from kinematic selections and imperfect
hardware, and incorporate those into the training and testing without biasing
the result.
We have also demonstrated this method on simulated multi-jet data in an
\atlas-like detector~\cite{lester2022hunting}.
By simulating parity-violating events from a Lorentz-violating extension
to the Standard Model, we show that this method build variables that are much
more sensitive than the $\alpha$ of Equation~\ref{eqn:cms_alpha} while
retaining its other attractive symmetries by encoding them into our function
architectures~\cite{lester2021stressed, tombs2021which}.

This work demonstrates that the `which is real?' method is ready to test for
parity violation in real LHC data.
Much like the emu work, however, its realization would require great diligence
to handle important considerations of systematic uncertainties and statistical
interpretations before it could be published through \atlas.

The `which is real?' method is general, and can test not only parity, but
symmetry under any transformation of the data, irrespective of whether it is
discrete or continuous.
In the context of modern machine learning, it is an example of
`self-supervised' learning, in which the training objective derives from the
data and not from additional labels~\cite{
Noroozi2016jigsaw,
multitaskself2017,
devlin2019bert
}, and quite similar to the `CLIP' objective, which also attempts to identify
a real examples from collections of alternatives~\cite{pmlr-v139-radford21a}.
What is special in our work is our application for scientific self-supervised
testing.

### 2022-09-28 main/frontmatter.tex
cut out preface junk; fix more atlas's; minor tweaks

\titlepage[of Homerton College]{%
A dissertation submitted to the University of Cambridge\\
for the degree of Doctor of Philosophy%
}

\begin{abstract}
\atlas\ is a detector experiment at \cern. $14~\eV[T]$.
\end{abstract}

\begin{declaration}
This dissertation is the result of my own work, except where explicit
reference is made to the work of others, and has not been submitted
for another qualification to this or any other university. This
dissertation does not exceed the word limit for the respective Degree
Committee.
\vspace*{1cm}
\begin{flushright}
Rupert Tombs
\end{flushright}
\end{declaration}

\begin{acknowledgements}
Of the many people who deserve thanks, some are particularly prominent,
such as my supervisor\dots
\\
This document uses parts of the ``hepthesis'' class~\cite{hepthesis}.
\end{acknowledgements}

@@

This thesis describes my research on various hoozywotsists.%
\footnote{test.}

\noindent
For this example, I'll just mention Chapter~\ref{chapter:searches}.
and Chapter~\ref{chapter:searches}..
\end{preface}

>>>>

\titlepage[Homerton College]{%
A dissertation submitted to the University of Cambridge\\
for the degree of Doctor of Philosophy%
}

\begin{abstract}
\TODO{searches for new effects in atlas are a core reason for its operations}

\TODO{this thesis presents such a search~\cite{atlas2022searches}}

\TODO{after briefly introducing its theoretical and experimental context}

\TODO{For given modelling of theory and experiment,
data analysis procedures dominate, and they take a very particular form
which we attempt to understand in its historical context
}

\TODO{
The result is compatible with many other observations,
that the Standard Model works very well at the LHC.
}

\TODO{supersymmetry is not falsifiable?}

\end{abstract}

\begin{declaration}
\noindent This dissertation is the result of my own work, except where explicit
reference is made to the work of others, and has not been submitted
for another qualification to this or any other university.
This dissertation does not exceed the word limit for the respective
Degree Committee.
\vspace*{1cm}
\begin{flushright}
Rupert~Tombs
\end{flushright}
\end{declaration}

\begin{acknowledgements}
<3
% Holly
% Dan
% Chris

% Friends who have come and gone from Homerton over the years

% Sarah Williams
% Ben Hooberman
% Jon Long

% Family, alive and dead.

This document uses parts of the ``hepthesis'' class~\cite{hepthesis}.
\end{acknowledgements}

@@

Data Intensive Science in High-Energy Physics

\TODO{Context of Standard Model theory, directions in which future developments
could improve it, and supersymmetric models}
Chapter~\ref{chapter:theory}

\TODO{the design of the \atlas\ detector}
Chapter~\ref{chapter:experiment}

\TODO{theory and practice of data science, as it could be for clarity
and how it is in our \atlas\ SUSY conventions}
Chapter~\ref{chapter:searches}

\TODO{the main event: the search with many of its gory details in addition
to those presented publically in the paper~\cite{atlas2022searches}}
Chapter~\ref{chapter:2ljets}

\paragraph{Other work.}
\TODO{Use some of the data analysis ideas of Chapter~\ref{chapter:searches}
to build predictive models
and test them to test for symmetries}
\TODO{citations}

\end{preface}
